\pkg{systemfit}: A Package for Estimating Systems of Simultaneous Equations in \proglang{R}

### 2008-09-12 paper/trunk/systemfit.Rnw
moved section "Using systemfit" (systemfit_usage.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@955 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\input{systemfit_usage.tex}
\input{systemfit_reliability.tex}
\input{systemfit_summary.tex}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Using systemfit]{Using \pkg{systemfit}}\label{sec:Usage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{prefix.string=trash/usage}
<<echo=FALSE>>=
options( prompt = "R> " )
@

In this section we demonstrate how to use the \pkg{systemfit} package.
First, we show the standard usage of \code{systemfit} by a simple example.
Second, several options that can be specified by the user are presented.
Then, the usage of \code{systemfit}
for a (classical) °seemingly unrelated regression° analysis
with panel data is described.
Finally, we demonstrate how to apply some statistical tests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[Standard usage of systemfit]{Standard usage of \code{systemfit}}
\label{sec:standard-usage}

As described in the previous section,
systems of equations can be econometrically estimated
with the function \code{systemfit}.
The only mandatory argument is \code{formula}.
Typically, it is a list of formulas to be estimated,
but it may also be a single formula for estimating a single-equation model.
Each formula is a standard regression formula in \proglang{R}
(see documentation of \code{formula}).

The following demonstration uses an example taken from
\citet[p.~685]{kmenta86}.
We want to estimate a small model of the US food market:
\begin{align}
\texttt{consump} & = \beta_1 + \beta_2 \cdot \texttt{price}
   + \beta_3 \cdot \texttt{income} \\
\texttt{consump} & = \beta_4 + \beta_5 \cdot \texttt{price}
   + \beta_6 \cdot \texttt{farmPrice} + \beta_7 \cdot \texttt{trend}
\end{align}
The first equation represents the demand side of the food market.
Variable \code{consump} (food consumption per capita) is the dependent variable.
The regressors are \code{price} (ratio of food prices to general consumer prices)
and \code{income} (disposable income) as well as a constant.
The second equation specifies the supply side of the food market.
Variable \code{consump} is the dependent variable of this equation as well.
The regressors are again \code{price} (ratio of food prices to general 
consumer prices) and a constant as well as 
\code{farmPrice} (ratio of preceding year's prices received by farmers
to general consumer prices) and 
\code{trend} (a time trend in years).
These equations can be estimated by OLS in \proglang{R} by
<<systemfit_usage>>=
library( "systemfit" )
data( "Kmenta" )
attach( Kmenta )
eqDemand <- consump ~ price + income
eqSupply <- consump ~ price + farmPrice + trend
eqSystem <- list( demand = eqDemand, supply = eqSupply )
fitols <- systemfit( eqSystem )
print( fitols )
@

The first line loads the \pkg{systemfit} package.
The second line loads example data that are included with the package.
They are attached to the \proglang{R} search path in line three.
In the fourth and fifth line, the demand and supply equations are specified,
respectively.%
\footnote{
A regression constant is always implied if not explicitly omitted.
}
In the sixth line, these equations are concatenated into a list and
are labeled \code{demand} and \code{supply}, respectively.%
\footnote{
If no labels are provided, the equations are numbered consecutively
( \code{eq1}, \code{eq2}, \ldots ).
}
Finally, in the last two lines, the regression is performed
and the estimation results are printed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[User options of systemfit]{User options of \code{systemfit}}
\label{sec:user-options}

The user can modify the default estimation method by providing
additional optional arguments,
e.g., to specify instrumental variables
or restrictions on the coefficients.
All optional arguments are described in the following:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Estimation method}

The optional argument \code{method} is a string
that determines the estimation method.
It must be either \code{"OLS"}, \code{"WLS"}, \code{"SUR"},
\code{"2SLS"}, \code{"W2SLS"}, or \code{"3SLS"}.
These methods correspond to the estimation methods
described in Sections~\ref{sec:Estimation-ols-wls-sur},
\ref{sec:Estimation-2sls-w2sls-3sls}, and~\ref{sec:Restrictions}.
The following command estimates the model described above as
°seemingly unrelated regression°.
<<systemfit_usage>>=
fitsur <- systemfit( eqSystem, method = "SUR" )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Instrumental variables}
The instruments for a 2SLS, W2SLS or 3SLS estimation can be specified
by the argument \code{inst}.
If the same instruments should be used for all equations,
\code{inst} must be a one-sided formula.%
\footnote{
A one-sided formula is a standard formula in \proglang{R}
without a dependent variable.}
If different instruments should be used for each equation,
\code{inst} must be a list that contains a one-sided formula
for each equation.
The following example uses instrumental variables
to estimate the model described above
by °three-stage least squares° (3SLS).
While the first command specifies the same instruments for all equations,
the second uses different instruments:
<<systemfit_usage>>=
fit3sls  <- systemfit( eqSystem, method = "3SLS",
   inst = ~ income + farmPrice + trend )
fit3sls2 <- systemfit( eqSystem, method = "3SLS",
   inst = list( ~ farmPrice + trend, ~ income + farmPrice + trend ) )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Data}
Having all data in the global environment or attached to the search path
is often inconvenient.
Therefore, \code{systemfit} has the argument \code{data}
to specify a data frame
that contains the variables of the model.
In the following example, we use this argument to specify
that the data for the estimation
should be taken from the data frame \code{Kmenta}.
Hence, we no longer need to attach this data frame before calling \code{systemfit}:
<<systemfit_usage>>=
fitsur <- systemfit( eqSystem, method = "SUR", data = Kmenta )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restrictions on the coefficients}
As outlined in Section~\ref{sec:Restrictions}, restrictions on the coefficients
can be specified in two ways.
One way is to use the matrix $R$ and the vector $q$
(see Section~\ref{sec:Restrictions}).
These restrictions can be specified symbolically
by argument \codeD{restrict}{matrix}
as in the generic function
\codeD{linear}{hypothesis} of the \pkg{car} package \citep{r-car-1.2-1, fox02a}.
This argument must be a vector of character strings,
where each element represents one linear restriction,
and each element must be either a linear combination of coefficients,
or a linear equation in the coefficients
(see documentation of function \codeD{linear}{hypothesis}
in the \pkg{car} package, \citealp{r-car-1.2-1, fox02a}).
We illustrate this by estimating the model under
the restriction $\beta_2 + \beta_6 = 0$.
Since the name of $\beta_2$
(coefficient of variable \code{price} in equation \code{demand})
is \code{demand\_price}
and the name of $\beta_6$
(coefficient of variable \code{farmPrice} in equation \code{supply})
is \code{supply\_farmPrice},
this restriction can be specified by
<<systemfit_usage,keep.source=TRUE>>=
restrict <- "demand_price + supply_farmPrice = 0"
fitsurRmat <- systemfit(eqSystem, method = "SUR",
   restrict.matrix = restrict)
@
\label{code:Rmat}

Alternatively, the restrictions via matrix $R$ and vector $q$
can be specified numerically.
The matrix $R$ can be specified with argument \codeD{restrict}{matrix}
and the vector $q$ with argument \codeD{restrict}{rhs}.
<<systemfit_usage,keep.source=TRUE>>=
Rmat <- matrix(0, nrow = 1, ncol = 7)
Rmat[1, 2] <- 1
Rmat[1, 6] <- 1
qvec <- c(0)
fitsurRmatNum <- systemfit(eqSystem, method = "SUR",
   restrict.matrix = Rmat, restrict.rhs = qvec)
@
The first line creates a $1 \times 7$ matrix of zeros,
where 1 is the number of restrictions and
7 is the number of unrestricted coefficients.
The following two lines specify this matrix in a way
that the multiplication with the coefficient vector
results in $ \beta_2 + \beta_6 $.
The fourth line creates a vector with a single element
that contains the right hand side of the restriction, i.e., zero.
Finally the coefficients are estimated under the restriction
$\beta_2 + \beta_6 = 0$.

The other way to specify restrictions on the coefficients is to modify the regressor matrix
by post-multiplying it with a matrix, say $M$
(see Section~\ref{sec:Restrictions}).
This kind of restriction can be specified
by setting argument \codeD{restrict}{regMat} equal to the matrix $M$.
We convert the restriction specified above to $\beta_2 = - \beta_6$,
and set $\beta_1 = \beta^\text{M}_1$, \ldots, $\beta_5 = \beta^\text{M}_5$,
$\beta_6 = - \beta^\text{M}_2$, and $\beta_7 =  \beta^\text{M}_6$.
We can do this in \proglang{R} by
<<systemfit_usage,keep.source=TRUE>>=
modRegMat <- matrix(0, nrow = 7, ncol = 6)
modRegMat[1:5, 1:5] <- diag(5)
modRegMat[6, 2] <- -1
modRegMat[7, 6] <- 1
fitsurRegMat <- systemfit(eqSystem, method = "SUR",
   restrict.regMat = modRegMat)
@
The first line creates a $7 \times 6$ matrix of zeros,
where 7 is the number of unrestricted coefficients and
6 is the number of restricted coefficients.
The following three lines specify the matrix $M$ (\code{modRegMat})
as described before.
Finally the coefficients are estimated under the restriction
$\beta^\text{M}_2 = \beta_2 = - \beta_6$.

Of course, the estimation results do not depend on the method
that was used to specify this restriction:
<<systemfit_usage>>=
all.equal( coef( fitsurRmat ), coef( fitsurRmatNum ) )
all.equal( coef( fitsurRmat ), coef( fitsurRegMat ) )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Iteration control}
The estimation methods WLS, SUR, W2SLS and 3SLS need a
covariance matrix of the residuals
that can be calculated from a first-step OLS or 2SLS estimation
(see Section~\ref{sec:residcov}).
This procedure can be iterated and at each iteration the covariance
matrix is calculated from the previous step estimation.
This iteration is repeated until the maximum number of iterations
is reached or the coefficient estimates have converged.
The maximum number of iterations is specified by argument \code{maxiter}.
Its default value is one, which means no iteration.
The convergence criterion is
\begin{equation}
   \sqrt{ \frac{ \sum_i (\beta_{i,g} - \beta_{i,g-1})^2 }
   { \sum_i \beta_{i,g-1}^2 }}
      < \texttt{tol} ,
\end{equation}
where $\beta_{i,g}$ is the $i$th coefficient of the $g$th iteration.
The default value of the convergence criterion (argument \code{tol})
is $10^{-5}$.

In the following example,
we estimate the model described above by iterated SUR:
<<systemfit_usage>>=
fitsurit <- systemfit( eqSystem, method = "SUR", maxiter = 500 )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Residual covariance matrix}
It was explained in Section~\ref{sec:residcov} that several different
methods have been proposed to calculate the residual
covariance matrix.
The user can specify, which method \code{systemfit} should use.
Possible values of the argument \code{methodResidCov} are presented in
Table~\ref{tab:methodResidCov}.
By default, \code{systemfit} uses Equation~\ref{eq:rcovGeomean}.

\begin{table}[H]
\centering
\begin{tabular}{|c|c|}
\hline
argument & equation \\
\code{methodResidCov} & \\
\hline
\code{"noDfCor"} & \ref{eq:rcovNoDfCor} \\
\hline
\code{"geomean"} & \ref{eq:rcovGeomean} \\
\hline
\code{"max"}     & \ref{eq:rcovMax} \\
\hline
\code{"Theil"}   & \ref{eq:rcovTheil} \\
\hline
\end{tabular}
\caption{Possible values of argument \code{methodResidCov}}
\label{tab:methodResidCov}
\end{table}

Furthermore, the user can specify
whether the means should be subtracted from the residuals
before Equations~\ref{eq:rcovNoDfCor}, \ref{eq:rcovGeomean}, \ref{eq:rcovMax},
or~\ref{eq:rcovTheil} are applied
to calculate the residual covariance matrix
(see Section~\ref{sec:residcov}).
The corresponding argument is called \code{centerResiduals}.
It must be either \code{TRUE} (subtract the means) or
\code{FALSE} (take the unmodified residuals).
The default value of \code{centerResiduals} is \code{FALSE}.

Moreover, if the coefficients are estimated under restrictions,
the user can use argument \code{residCovRestricted} to specify
whether the residual covariance matrix
for a WLS, SUR, W2SLS, or 3SLS estimation
should be obtained from a restricted or from an unrestricted
first-step estimation
(see Section~\ref{sec:residcov}).
If this argument is \code{TRUE} (the default),
the residual covariance matrix is obtained from a restricted
OLS or 2SLS estimation.
If it is \code{FALSE},
the residual covariance matrix is obtained from an unrestricted
first-step estimation.

Finally, argument \code{residCovWeighted} can be used to decide,
whether the residual covariance matrix for a SUR (3SLS) estimation
should be obtained from a WLS (W2SLS) estimation
instead of from an OLS (2SLS) estimation
(see Section~\ref{sec:residcov}).
By default, \code{residCovWeighted} is \code{FALSE},
which means that the residuals of an OLS (2SLS) estimation are used
to compute the residual covariance matrix.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{3SLS formula}
As discussed in Sections~\ref{sec:Estimation-2sls-w2sls-3sls} and~\ref{sec:Restrictions},
there exist several different
methods to perform a 3SLS estimation.
The user can specify the method by argument \code{method3sls}.
Possible values are presented in Table~\ref{tab:method3sls}.
The default value is \code{"GLS"}.

\begin{table}[H]
\centering
\begin{tabular}{|c|c|c|}
\hline
argument           & equation       & equation \\
\code{method3sls} & (unrestricted) & (restricted) \\
\hline
\code{"GLS"}     & \ref{eq:2sls-w2sls-3sls} & \ref{eq:2sls-w2sls-3sls-r} \\
\hline
\code{"IV"}      & \ref{eq:3slsIv}      & \ref{eq:3slsIvR} \\
\hline
\code{"GMM"}     & \ref{eq:3slsGmm}     & \ref{eq:3slsGmmR} \\
\hline
\code{"Schmidt"} & \ref{eq:3slsSchmidt} & \ref{eq:3slsSchmidtR} \\
\hline
\code{"EViews"}  & \ref{eq:3slsEViews}  & \ref{eq:3slsEViewsR} \\
\hline
\end{tabular}
\caption{Possible values of argument \code{method3sls}}
\label{tab:method3sls}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Sigma squared}
In case of OLS or 2SLS estimations, argument \code{singleEqSigma}
can be used to specify,
whether different $\sigma^2$s for each single equation
or the same $\sigma^2$ for all equations should be used.
If argument \code{singleEqSigma} is \code{TRUE},
$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} or~\ref{eq:cov-2sls-w2sls-3sls}
is set to $\SHat \otimes I_T$.
In contrast, if argument \code{singleEqSigma} is \code{FALSE},
$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} or~\ref{eq:cov-2sls-w2sls-3sls}
is set to $\sHat^2 I_{G \cdot T}$.
In case of an unrestricted regression,
argument \code{singleEqSigma} is \code{TRUE} by default.
However, if the coefficients are estimated under restrictions,
this argument is \code{FALSE} by default.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{System options}
Furthermore, two options regarding some internal calculations are available.
First, argument \code{solvetol} specifies the tolerance level
for detecting linear dependencies when inverting a matrix or
calculating a determinant (using functions \code{solve} and \code{det}).
The default value depends on the used computer system and is equal to the
default tolerance level of \code{solve} and \code{det}.

Second, argument \code{useMatrix} specifies
whether the \pkg{Matrix} package \citep{r-matrix-07}
should be used for all computations where matrices are involved
(see Section~\ref{sec:code-efficiency}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Returned data objects}

Finally, the user can decide
whether \code{systemfit} should return some data objects.
Argument \code{model} indicates
whether a data frame with the data of the model
should be returned.
Its default value is \code{TRUE},
i.e., the model frame is returned.
Arguments \code{x}, \code{y},
and \code{z} specify
whether the model matrices ($X_i$),
the responses ($y_i$), and the matrices of instrumental variables ($Z_i$),
respectively, should be returned.
These three arguments are \code{FALSE} by default,
i.e., these data objects are not returned.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[Summary results with summary.systemfit]
   {Summary results with \code{summary.systemfit}}

The \code{summary} method can be used
to compute and print summary results of objects
returned by \code{systemfit}.
<<systemfit_usage>>=
summary( fitsur )
@
First, the estimation method is reported
and a few summary statistics for the entire system
and for each equation are given.
Then, the covariance matrix used for estimation
and the covariance matrix as well as the correlation matrix
of the (final) residuals are printed.
Finally, the estimation results of each equation are reported:
the formula of the estimated equation,
the estimated coefficients, their standard errors, $t$ values, $P$ values
and codes indicating their statistical significance,
as well as some other statistics like
the standard error of the residuals and
the $R^2$ value of the equation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection[Degrees of freedom for t tests]
   {Degrees of freedom for $t$ tests}

The \code{summary} method for \code{systemfit} objects has an optional argument \code{useDfSys}.
It selects the approach
that is applied by \code{systemfit}
to determine the degrees of freedom of $t$ tests of the estimated coefficients
(Section~\ref{sec:degreesOfFreedom}).
If argument \code{useDfSys} is \code{TRUE},
the degrees of freedom of the whole system are taken.
In contrast, if \code{useDfSys} is \code{FALSE},
the degrees of freedom of the single equation are taken.
If the coefficients are estimated under restrictions,
argument \code{useDfSys} is \code{TRUE} by default.
However, if no restrictions on the coefficient are specified,
the default value of \code{useDfSys} is \code{FALSE}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Reduce amount of printed output}

The optional arguments \code{residCov} and \code{equations}
can be used reduce the amount of the printed output.
Argument \code{residCov} specifies
whether the covariance matrix and
the correlation matrix
of the residuals are printed.
Argument \code{equations} specifies
whether summary results of each equation are printed.
By default, both arguments are \code{TRUE}.
The following command returns a sparse summary output:
<<systemfit_usage>>=
summary( fitsur, residCov = FALSE, equations = FALSE )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Panel data}

The \code{systemfit} function can also be used
for a (classical) °seemingly unrelated regression° analysis
with panel data.
For this type of analysis,
the data must be provided in a transformed data frame of class \codeD{plm}{dim}%
\footnote{
Generally, panel data can be either in °long format°
(different individuals are arranged below each other)
or in °wide format°
(different individuals are arranged next to each other).
For this analysis, the data must be in °long format°.
}
which can be created with the function \codeD{plm}{data}
from the \proglang{R} package \pkg{plm}
\citep{r-plm-0.3-1}.
In contrast to the previously described usage of \code{systemfit},
argument \code{formula} must be a single equation (object of class \code{formula}).
This formula is estimated for all individuals.

We demonstrate the application of \code{systemfit}
to panel data
using an example taken from \citet[p.~340]{greene03}
that is based on \citet{grunfeld58}.
We want to estimate a model for gross investment of 5 US firms
in the years 1935--1954:
\begin{equation}
\texttt{invest}_{it} = \beta_1 + \beta_2 \cdot \texttt{value}_{it} +
   \beta_3 \cdot \texttt{capital}_{it}
\end{equation}
where \code{invest} is the gross investment of firm $i$ in year $t$,
\code{value} is the market value of the firm at the end of the
previous year, and
\code{capital} is the capital stock of the firm at the end of the
previous year.

This model can be estimated by
<<systemfit_usage>>=
data( "GrunfeldGreene" )
library( "plm" )
GGPanel <- plm.data( GrunfeldGreene, c( "firm", "year" ) )
greeneSur <- systemfit( invest ~ value + capital, method = "SUR",
   data = GGPanel )
@
The first line loads the example data set \code{GrunfeldGreene}
that is included in the \pkg{systemfit} package.
The second line loads the \pkg{plm} package
and the following line specifies a data frame of class \codeD{plm}{dim},
where the variables \code{firm} and \code{year} indicate
the individual (cross-section) and time identifier, respectively.
Finally, a seemingly unrelated regression is performed.

The optional argument \code{pooled} is a logical variable indicating
whether the coefficients are restricted to be equal for all
individuals.
By default, this argument is set to \code{FALSE}.
The following command does a seemingly unrelated regression
of the same model as before,
but with coefficients restricted to be equal for all individuals.
<<systemfit_usage>>=
greeneSurPooled <- systemfit( invest ~ value + capital, method = "SUR",
   data = GGPanel, pooled = TRUE )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing linear restrictions}

As described in Section~\ref{sec:testingRestrictions},
linear restrictions can be tested by an $F$ test, two Wald tests and
an LR test.
The \pkg{systemfit} package provides
the method \codeD{linear}{hypothesis} for the $F$ and Wald tests
as well as the method \code{lrtest} for LR tests.

We will now test the restriction $\beta_2 = -\beta_6$
that was specified by the matrix \code{Rmat} and the vector \code{qvec}
in the example above (p.~\pageref{code:Rmat}).
<<systemfit_usage>>=
linear.hypothesis( fitsur, Rmat, qvec, test = "FT" )

linear.hypothesis( fitsur, Rmat, qvec, test = "F" )

linear.hypothesis( fitsur, Rmat, qvec, test = "Chisq" )

lrtest( fitsurRmat, fitsur )
@
The linear restrictions are tested by \citeauthor{theil71}'s $F$ test
(Equation~\ref{eq:f-test-theil}) first,
second by the $F$ statistic of a Wald test
(Equation~\ref{eq:f-test-wald}),
third by the $\chi^2$ statistic of a Wald test
(Equation~\ref{eq:chi2-test-wald}),
and finally by an LR test
(Equation~\ref{eq:lr-test}).

The first argument of the \codeD{linear}{hypothesis} method for \code{systemfit} objects
must be an unrestricted regression
returned by \code{systemfit}.
The second and third arguments are the restriction matrix $R$ and
the optional vector $q$, as described in Section~\ref{sec:Restrictions}.
Analogously to the argument \codeD{restrict}{matrix}
of the \code{systemfit} function,
the restrictions can be specified either in matrix form or symbolically.
The optional argument \code{test} must be
a character string, \code{"FT"}, \code{"F"}, or \code{"Chisq"},
specifying whether to compute
\citeauthor{theil71}'s finite-sample $F$ test
(with approximate $F$ distribution)
the finite-sample Wald test (with approximate $F$ distribution)
or the large-sample Wald test (with asymptotic $\chi^2$ distribution).

All arguments of the \code{lrtest} method for \code{systemfit} objects
must be fitted model objects returned by \code{systemfit}.
It consecutively compares all provided fitted model objects.

All tests print a short description of the test
and the tested model objects first.
Then, a small table is printed,
where each row belongs to one (unrestricted or restricted) model.
The second row reports (amongst others)
the degree(s) of freedom of the test,
the empirical test statistic,
and the marginal level of significance ($P$ value).
Although all tests check the same hypothesis,
there is some variation of the $P$ values.
However, all tests suggest the same decision:
The null hypothesis $\beta_2 = -\beta_6$ cannot be rejected at any
reasonable level of significance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hausman test}
A Hausman test, which is described in Section~\ref{sec:hausman},
can be carried out with following commands:
<<systemfit_usage>>=
fit2sls  <- systemfit( eqSystem, method = "2SLS",
   inst = ~ income + farmPrice + trend, data = Kmenta )
fit3sls <- systemfit( eqSystem, method = "3SLS",
   inst = ~ income + farmPrice + trend, data = Kmenta )
hausman.systemfit( fit2sls, fit3sls )
@
<<echo=FALSE>>=
hausmantest <- hausman.systemfit( fit2sls, fit3sls )
@
First of all, the model is estimated by 2SLS and then by 3SLS.
Finally, in the last line
the test is carried out by the command \codeD{hausman}{systemfit}.
This function requires two arguments:
the result of a 2SLS estimation and
the result of a 3SLS estimation.
The Hausman test statistic is
\Sexpr{round( hausmantest$statistic, digits = 3)},
which has a $\chi^2$ distribution
with \Sexpr{hausmantest$parameter} degrees of freedom
under the null hypothesis.
The corresponding $P$ value is
\Sexpr{round( hausmantest$p.value, digits = 3 )}.
This shows that the null hypothesis is not rejected
at any reasonable level of significance.
Hence, we can assume that the 3SLS estimator is consistent.

\input{systemfit_reliability.tex}
\input{systemfit_summary.tex}
